Comparing variability across fMRI analysis software packages and numerical errors

### 2021-12-09 paper/paper.tex
Address a few comments

We measured between-tool (BT) variability by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. These software versions were identical to the study
in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation \gk{Did we verify that single-threading generally didn't lead to different results than 7-threading? As in, were they all ostensibly sampled from the same random distribution?}. All
the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake~@~2.4~GHz
 (40 cores/node) CPU and 8~GB of RAM per core. To facilitate portability and reproducibility,
 we encapsulated the
above-mentioned software packages in Docker\gk{CC doesn't support Docker, so you mean Singularity?} container images based on CentOS 7.3.

@@

Table~\ref{table:pipeline-stats} presents summary statistics for the
group-level t-statistics. For each tool pair (A, B), BT variability was
significantly larger than machine error in tool A or B, in both thresholded
and unthresholded maps (Wilcoxon signed-rank test p~\textless~$10^{-162}$ \gk{we don't have sufficient sample size to say that p is that small, so we should probably say less than $10^{-5}$ or something else we could actually arrive at with our number of repetitions, pipeline pairs, and samples.}
for all tests, t-test p=0 \gk{similar comment re p value} for all tests).
Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B},\textbf{C} shows that
these global differences were confirmed regionally. 
In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's R
in [0.56, 0.58], p=0.0, Figure~\ref{fig:unthresh-maps}-\textbf{D}). Interestingly, the correlation 
appears driven by a set of voxels exhibiting comparable BT and
WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have 
any spatial consistency \tristan{Ali, check for intensity differences between these voxels and the other ones}.

\setlength{\tabcolsep}{5pt}
\begin{table}[h]
    \centering
    \begin{tabular}{cccccc|cc}
        \toprule
        && \multicolumn{4}{c|}{Group map} & \multicolumn{2}{c}{Subject maps}\\
        \multirow{2}{*}{}& {} & \multicolumn{2}{c}{Thresholded} & \multicolumn{2}{c|}{Unthresholded} & \multicolumn{2}{c}{Unthresholded} \\
        % \cmidrule{3-8}
        {} & {} & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ \\
        \midrule
        \rowcolor{lightgray!50}
        {Between Tools} & FSL vs. SPM        &  1.282       & 0.525      & 0.443     & 0.344  & 0.366       & 0.293     \\
        \rowcolor{lightgray!50}
        {(BT)} & FSL vs. AFNI                &  1.548       & 0.616      & 0.547     & 0.441  & 0.439       & 0.352     \\
        \rowcolor{lightgray!50} 
        {} & AFNI vs. SPM                    &  1.475       & 0.672      & 0.608     & 0.477  & 0.491       & 0.381     \\
        {Within Tool} & FSL                  &  0.354       & 0.491      & 0.082     & 0.065  & 0.077       & 0.054     \\
        {(WT)}   & SPM                       &  0.252       & 0.448      & 0.054     & 0.045  & 0.048       & 0.037     \\
        {}   & AFNI                          &  0.434       & 0.524      & 0.128     & 0.135  & 0.108       & 0.131     \\
        \bottomrule
    \end{tabular}
    \caption{Voxel-wise mean and standard deviation of BT and WT variability
    in t-statistics maps. \tristan{add stats on subject-level thresholded maps}}
    \label{table:pipeline-stats}
\end{table}

@@

  Table~\ref{table:pipeline-stats} also presents summary statistics for
  subject-level unthresholded t-statistics maps. As for group-level maps,
  BT variability was consistently larger than WT variability (Wilcoxon
  signed-rank test p~\textless~$10^{-160}$\gk{pval comment} for all tests, t-test p=0 for
  all tests). However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
  some subjects, WT variability approached and even
  surpassed BT in some regions. As for the group maps, BT and WT
  variability appeared moderately correlated for all subjects (R in [0.442,
  0.612], p=0.0 for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).

@@

While the previous results were obtained for machine error, we also
evaluated numerical variability across different virtual precisions for FSL
and found that the virtual precision of t=17 bits minimized the RMSE
between BT and WT in unthresholded group maps. Figure~\ref{fig:gnp-mni} shows BT
and WT in the corresponding group maps obtained at t=17 bits, showing that
BT and WT reach comparable magnitudes in some regions. BT and WT remained
moderately correlated at this precision. \tristan{In the text, add statistics similar to Table 2: is BT still larger than WT here?}

>>>>

We measured between-tool (BT) variability by running the pipelines
described in~\cite{bowring2019exploring} with FSL version 5.0.10, AFNI
version 18.1.09, and SPM12 version r7771 executed with GNU/Octave version
5.2. These software versions were identical to the study
in~\cite{bowring2019exploring} except for SPM for which we used GNU/Octave
instead of MATLAB to enable mathematical function instrumentation using
Fuzzy Libmath. Indeed, MATLAB uses its own built-in mathematical functions,
which prevents the use of Fuzzy Libmath. In AFNI, we set the number of
threads to 7 even though AFNI executions
in~\cite{bowring2019exploring} were single-threaded. This was meant to
reduce the time overhead resulting from Fuzzy Libmath instrumentation 
\gk{Did we verify that single-threading generally didn't lead to different results than 7-threading? As in, were they all ostensibly sampled from the same random distribution?}. 
\ali{We verified that multithreading leads to different results (e.g. single-threading vs, 7-threading). 
However, single-threaded executions couldn't reproduce the original results.}
All the analyses were conducted on the CentOS 7.3 operating system. The
computations were performed on \href{https://www.computecanada.ca}{Compute
Canada's} Béluga cluster nodes, each with 2$\times$ Intel Gold 6148 Skylake~@~2.4~GHz
 (40 cores/node) CPU and 8~GB of RAM per core. To facilitate portability and reproducibility,
 we encapsulated the
above-mentioned software packages in Docker container images based on CentOS 7.3.
Although, we converted Docker containers to Singularity images to enable running on cluster nodes.

@@

Table~\ref{table:pipeline-stats} presents summary statistics for the
group-level t-statistics. For each tool pair (A, B), BT variability was
significantly larger than machine error in tool A or B, in both thresholded
and unthresholded maps (Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
% \gk{we don't have sufficient sample size to say that p is that small,
% so we should probably say less than $10^{-5}$ or something else we could actually
% arrive at with our number of repetitions, pipeline pairs, and samples.}
% for all tests, t-test p=0 \gk{similar comment re p value} for all tests).
Figure~\ref{fig:unthresh-maps}-\textbf{A},\textbf{B},\textbf{C} shows that
these global differences were confirmed regionally. 
In addition, BT and WT variability appeared moderately correlated across voxels (Pearson's R
in [0.56, 0.58], p=0.0, Figure~\ref{fig:unthresh-maps}-\textbf{D}). Interestingly, the correlation 
appears driven by a set of voxels exhibiting comparable BT and
WT values located around the identity line in Figure~\ref{fig:unthresh-maps}D, which, however, did not seem to have 
any spatial consistency.

\setlength{\tabcolsep}{5pt}
\begin{table}[h]
    \centering
    \begin{tabular}{cccccc|cc}
        \toprule
        && \multicolumn{4}{c|}{Group map} & \multicolumn{2}{c}{Subject maps}\\
        \multirow{2}{*}{}& {} & \multicolumn{2}{c}{Thresholded} & \multicolumn{2}{c|}{Unthresholded} & \multicolumn{2}{c}{Unthresholded} \\
        % \cmidrule{3-8}
        {} & {} & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ & $\mu$ & $\sigma$ \\
        \midrule
        \rowcolor{lightgray!50}
        {Between Tools} & FSL vs. SPM        &  1.282       & 0.525      & 0.443     & 0.344  & 0.366       & 0.293     \\
        \rowcolor{lightgray!50}
        {(BT)} & FSL vs. AFNI                &  1.548       & 0.616      & 0.547     & 0.441  & 0.439       & 0.352     \\
        \rowcolor{lightgray!50} 
        {} & AFNI vs. SPM                    &  1.475       & 0.672      & 0.608     & 0.477  & 0.491       & 0.381     \\
        {Within Tool} & FSL                  &  0.354       & 0.491      & 0.082     & 0.065  & 0.077       & 0.054     \\
        {(WT)}   & SPM                       &  0.252       & 0.448      & 0.054     & 0.045  & 0.048       & 0.037     \\
        {}   & AFNI                          &  0.434       & 0.524      & 0.128     & 0.135  & 0.108       & 0.131     \\
        \bottomrule
    \end{tabular}
    \caption{Voxel-wise mean and standard deviation of BT and WT variability
    in t-statistics maps.}
    \label{table:pipeline-stats}
\end{table}

@@

  Table~\ref{table:pipeline-stats} also presents summary statistics for
  subject-level unthresholded t-statistics maps. As for group-level maps,
  BT variability was consistently larger than WT variability (Wilcoxon
  signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
  However, Figure~\ref{fig:unthresh-maps-sbj} shows that for
  some subjects, WT variability approached and even
  surpassed BT in some regions. As for the group maps, BT and WT
  variability appeared moderately correlated for all subjects (R in [0.442,
  0.612], p=0.0 for all subjects, see Supplemental Material~\ref{sec:supp-subjects}).

@@

While the previous results were obtained for machine error, we also
evaluated numerical variability across different virtual precisions for FSL
and found that the virtual precision of t=17 bits minimized the RMSE
between BT and WT in unthresholded group maps.
FSL produced group maps with $\mu=0.325$ and $\sigma=0.287$ at this virtual precision,
which indicates a bit lower variability compared to BT variability
(Wilcoxon signed-rank test and t-test p~\textless~$10^{-5}$ for all tests).
Figure~\ref{fig:gnp-mni} shows BT
and WT in the corresponding group maps obtained at t=17 bits, showing that
BT and WT reach comparable magnitudes in some regions. BT and WT remained
moderately correlated at this precision.
